\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-06-24 syn-translate.tex
Save syn-translate

\newcommand{\denotesT}[1]{{{\llbracket {#1} \rrbracket}_T}}
\newcommand{\Sigr}[2]{{ "Sig"_r~{#1}~{#2} }}

We use $\denotesT{}$ as the translation. This translation is invariant under the subpart. Since we are in an intrinsic setting, everything here is well-typed (well-formed)including this translation. Thus $\denotesS{T}$ means the same as $\denotesS{\goodType{\Gamma}{T}{}}$ (i.e. every type is tracking its context). In other words, every derivation of any judgement will track its ``context''. This will be clearer in the (pseudo-)Agda formalization.

\begin{align*}
  \goodType{\Gamma}{\Sigr{\Gamma}{n}}{} & \text{ when } \goodCtx{\Gamma}{} \text{ and } n \in \mathbb{N} \\
  \denotesT{\goodSig{\Gamma}{\sigma}{n}},&\denotesT{\Sigr{\Gamma}{n}}, \text{ and } \denotesT{\cC} \text{ are mutually recursively defined together, } \\
  & \text{and they define inductively on the signature length} \\  
  \denotesT{\goodSig{\Gamma}{\sigma}{n}} &= \goodTerm{\Gamma}{\sigma}{\Sigr{\Gamma}{n}} \\ 
  \denotesT{\Sigr{\Gamma}{n+1}} &= \Sigma~(\Sigr{\Gamma}{n})~(\Sigma)
\end{align*}

>>>>

\newcommand{\denotesT}[1]{{{\llbracket {#1} \rrbracket}_T}}
\newcommand{\Sigr}[2]{{ "Sig"^r~{#1}~{#2} }}

We use $\denotesT{}$ as the translation. This translation is invariant under the subpart. Since we are in an intrinsic setting, everything here is well-typed (well-formed)including this translation. Thus $\denotesT{T}$ means the same as $\denotesT{\goodType{\Gamma}{T}{}}$ (i.e. every type is tracking its context). In other words, every derivation of any judgement will track its ``context''. This will be clearer in the (pseudo-)Agda formalization.

\begin{align*}
  \text{We will define a new type }& \goodType{\Gamma}{\Sigr{\Gamma}{n}}{} \text{ when } \goodCtx{\Gamma}{} \text{ and } n \in \mathbb{N} \\
  \denotesT{\goodSig{\Gamma}{\_}{n}},\ &({\Sigr{\Gamma}{n}}), \text{ and } \ \denotesT{\cC} \text{ are mutually recursively defined together, } \\
  & \text{and they define inductively on the signature length} \\  
  \denotesT{\goodSig{\Gamma}{\_}{n}} &= (\goodTerm{\Gamma}{\_}{\Sigr{\Gamma}{n}}) \text{, s.t. } \denotesT{\goodSig{\Gamma}{\sigma}{n}} \iff \goodTerm{\Gamma}{\sigma}{\Sigr{\Gamma}{n}} \\ 
  {\Sigr{\Gamma}{(n+1)}} &= 
    \Sigma~(\Sigr{\Gamma}{n})
          ~(\Sigma~\cU
                  ~(\Sigma~(\Pi(\cC (\pi_2[\pi_1]))(El~\pi_2[\pi_1]))~\cU)) \\
  \Sigr{\Gamma}{0} &= \top \\
  \denotesT{\goodType{\Gamma}{\cC \sigma}{}} &\text{ defines upon } \denotesT{\sigma} \\
  \denotesT{\goodType{\Gamma}{\cC \sigma}{}} &= 
    \Sigma~(\denotesT{\cC ("pjl"~\sigma)})~(El \ ("app"~("pjr"^3~\sigma))[(\pi_1, "app"~("pjl"~("pjr"^2~t)))]) \\
      &\text{when } \goodSig{\Gamma}{\sigma}{n+1} \\
  \denotesT{\goodType{\Gamma}{\cC \sigma}{}} &= \top \quad
      \text{when } \goodSig{\Gamma}{\sigma}{0} \\ 
  \denotesT{\goodType{\Gamma}{\cL\sigma}{}} & \text{ is defined upon } \denotesT{\sigma} \\
  & \text{ and inductively on the signature length} \\ 
  \denotesT{\goodType{\Gamma}{\cL\sigma}{}} &=
  \denotesT{\cL("pjl"~\sigma)} \times \Pi(El~("pjl"~("pjr"~\sigma)))(El~("app"~("pjr"^3~\sigma))) \\
  &\text{when } \goodSig{\Gamma}{\sigma}{n+1} \\
  \denotesT{\goodType{\Gamma}{\cL\sigma}{}} &= \top \quad \text{when } \goodSig{\Gamma}{\sigma}{0} \\
  \denotesT{\goodSig{\Gamma}{\nu^+~\sigma~f~T}{n+1}} &= (\sigma, "c"~A, "lam"~f, "lam"~("c"~T)) \\ 
  \denotesT{\goodTerm{\Gamma}{\mu^+~m~t}{\cL(\nu^+~\sigma~T)}} &= (m, "lam"~t) \\ 
  \denotesT{\goodTerm{\Gamma}{\cCt m}{\cC\sigma}} & \text{ is defined upon } \denotesT{m}\\
  & \text{ and inductively on the signature length} \\ 
  \denotesT{\goodTerm{\Gamma}{\cCt m}{\cC\sigma}} &= 
  (\denotesT{\cCt o}, t["sf"~f][("id", \denotesT{\cCt o})]) \\
  & \text{when } \goodSig{\Gamma}{\sigma}{n+1}, \text{ and } o = "pjl"~m, t = "app"~("pjr"~m), f = p_f\nu~\sigma \\ 
  \denotesT{\goodTerm{\Gamma}{\cCt m}{\cC \sigma}} &= () \quad \text{when } \goodSig{\Gamma}{\sigma}{0}
\end{align*}

### 2022-06-24 syntax.tex
Finishing Language Design Section. Try to fix graph embedding and format.
Save syn-translate

\begin{figure}[H]
  \begin{minipage}[b]{0.3\linewidth}
      $$
      \Rule[name=Tm]
      {\goodCtx{\Gamma}{i} \quad \goodType{\Gamma}{T}{j}}
      {\goodTerm{\Gamma}{\_}{T}}
      $$
      $$
      \Rule[name=Tms]
      {\goodCtx{\Gamma}{i} \quad \goodCtx{\Delta}{i}}
      {\goodSub{\Gamma}{\_}{\Delta} \text{ is well-formed}}
      $$
      $$
      \Rule[name=Sig]
      {\goodCtx{\Gamma}{i} \quad n \in \nat}
      {\goodSig{\Gamma}{\_}{n} \text{ is well-formed}}
      $$
      $$
      \Rule[name=Seal]
      {\goodCtx{\Gamma}{i} \quad \goodSig{\Gamma}{\sigma}{n} 
      \quad \goodType{\Gamma}{A}{} }
      {\goodSeal{\Gamma}{\_}{\sigma}{A} \text{ is well-formed}}
      $$
  \end{minipage}
  \begin{minipage}[b]{0.6\linewidth}
    \begin{minted}[]{agda}
      data Con   : Set 
      data Ty    : Con → Set   
      data Tms   : Con → Con → Set 
      data Tm    : (Γ : Con) → Ty Γ  → Set 
      data Sig  : Con → ℕ → Set
      data WSig : Con → ℕ → Set 
      data Inh  : (Γ : Con) → Sig Γ n → Sig Γ m → Set
      Seal : (Γ : Con) → Sig Γ n → Sig Γ n → Set
    \end{minted}
  \end{minipage}
\begin{align*}
  &\Gamma~:~"Con"&&\mathboxtext{\goodCtx{\Gamma}{}} &&T~:~"Ty"~\Gamma&&\mathboxtext{\goodType{\Gamma}{T}{}}  &&\sigma~:~"Tms"~\Gamma~\Delta&&\mathboxtext{\goodSub{\Gamma}{\sigma}{\Delta}}  \\ & t~:~"Tm"~\Gamma~T&&\mathboxtext{\goodTerm{\Gamma}{t}{T}} 
  &&\sigma~:~"Sig"~\Gamma~n&&\mathboxtext{\goodSig{\Gamma}{\sigma}{n}}  &&\sigma~:~"WSig"~\Gamma~n&&\mathboxtext{\goodWSig{\Gamma}{\sigma}{n}} \\ &h~:~"Inh"~\Gamma~\sigma~\tau&&\mathboxtext{\goodInh{\Gamma}{h}{\sigma}{\tau}} && f~:~"Seal"~\Gamma~\sigma~A&&\mathboxtext{\goodSeal{\Gamma}{f}{\sigma}{A}}
\end{align*}
\caption{Well-formedness of Four New Judgements, and Agda-encoding-Judgements correspondence}
\end{figure}

@@

$$
\Rule[name=Compile]
{}
{\goodType{\Gamma}{\cC \nu\cdot \equiv \top}{} 
\\ \goodTerm{\Gamma}{ \cCt \mu\cdot \equiv ()}{\cL \nu\cdot}
\\
\goodType{\Gamma}{\cC (\nu^+ \ \sigma \ \{f\} \ T) \equiv 
    \Sigma (\cC \sigma) T[sf \ f]}{}
\\\\ \goodTerm{\Gamma}{\cCt \ (\mu^+ \ o \ \{f\}\ t) \equiv ((\cCt \ o), t[sf \ f][(id, \cCt \ o)]) }{}
}
$$
% $$
% \Rule
% { \goodSeal{\Gamma}{f}{\sigma}{A}
% \\ \goodType{\Gamma, A}{T}{}
% \\ \goodTerm{\Gamma}{t}{T}
% }{}
% $$

@@

\judgebox{[\goodSeal{\Gamma}{s}{\sigma}{A}] 
= [\goodTerm{\Gamma, \cC\sigma}{s}{A[\pi_1]}] }
$$
\Rule[name=Seal Subst]
{\goodSeal{\Delta}{s}{\sigma}{A}
  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
{\goodSeal{\Gamma}{s[\gamma]}{\sigma[\gamma]}{A[\gamma]}}
=
\frac
{\goodTerm{\Delta, \cC\sigma}{s}{A[\pi_1]}
  \quad  \goodSub{\Gamma}{\gamma}{\Delta}  }
{\goodTerm{\Gamma,\cC\sigma[\gamma]}{s[\gamma^\uparrow]}{A[\pi_1][\gamma^\uparrow]\equiv A[\gamma][\pi_1]}}
$$
$$
\Rule[name=Seal-Id]
{}
{\goodSeal{\Gamma}{id_s}{\sigma}{\cC \sigma}}
= \goodTerm{\Gamma, \cC\sigma}{\pi_2}{\cC\sigma[\pi_1]}
$$
\begin{figure}[H]

@@

And these are also the cases we don't need to seal anything (i.e. doing exhaustiveness checking when using recursor), and thus the "Seal-Id" is used. For notation clearance, we will sometime use $sf'$ to emphasize that we consider that seal $s$ as a term $(sf' \ s)$.

>>>>

\begin{figure}[!htb]
  \begin{align*}
    &\Gamma~:~"Con"&&\mathboxtext{\goodCtx{\Gamma}{}} &&T~:~"Ty"~\Gamma&&\mathboxtext{\goodType{\Gamma}{T}{}}  &&\sigma~:~"Tms"~\Gamma~\Delta&&\mathboxtext{\goodSub{\Gamma}{\sigma}{\Delta}}  \\ & t~:~"Tm"~\Gamma~T&&\mathboxtext{\goodTerm{\Gamma}{t}{T}} 
    &&\sigma~:~"Sig"~\Gamma~n&&\mathboxtext{\goodSig{\Gamma}{\sigma}{n}}  &&\sigma~:~"WSig"~\Gamma~n&&\mathboxtext{\goodWSig{\Gamma}{\sigma}{n}} \\ &h~:~"Inh"~\Gamma~\sigma~\tau&&\mathboxtext{\goodInh{\Gamma}{h}{\sigma}{\tau}} && f~:~"Seal"~\Gamma~\sigma~A&&\mathboxtext{\goodSeal{\Gamma}{f}{\sigma}{A}}
  \end{align*}
\caption{Judgements and their Agda Encoding}
\end{figure}

\begin{figure}[!htb]
  \begin{minipage}[b]{0.3\linewidth}
      $$
      \Rule[name=Tm]
      {\goodCtx{\Gamma}{i} \quad \goodType{\Gamma}{T}{j}}
      {\goodTerm{\Gamma}{\_}{T}}
      $$
      $$
      \Rule[name=Tms]
      {\goodCtx{\Gamma}{i} \quad \goodCtx{\Delta}{i}}
      {\goodSub{\Gamma}{\_}{\Delta} \text{ is well-formed}}
      $$
      $$
      \Rule[name=Sig]
      {\goodCtx{\Gamma}{i} \quad n \in \nat}
      {\goodSig{\Gamma}{\_}{n} \text{ is well-formed}}
      $$
      $$
      \Rule[name=Seal]
      {\goodCtx{\Gamma}{i} \quad \goodSig{\Gamma}{\sigma}{n} 
      \quad \goodType{\Gamma}{A}{} }
      {\goodSeal{\Gamma}{\_}{\sigma}{A} \text{ is well-formed}}
      $$
  \end{minipage}
  \begin{minipage}[b]{0.6\linewidth}
    \begin{minted}[]{agda}
      data Con   : Set 
      data Ty    : Con → Set   
      data Tms   : Con → Con → Set 
      data Tm    : (Γ : Con) → Ty Γ  → Set 
      data Sig  : Con → ℕ → Set
      data WSig : Con → ℕ → Set 
      data Inh  : (Γ : Con) → Sig Γ n → Sig Γ m → Set
      Seal : (Γ : Con) → Sig Γ n → Sig Γ n → Set
    \end{minted}
  \end{minipage}

\caption{Well-formedness of Four New Judgements}
\end{figure}

@@

$$
\Rule[name=Compile]
{}
{\goodType{\Gamma}{\cC \nu\cdot \equiv \top}{} 
\\ \goodTerm{\Gamma}{ \cCt \mu\cdot \equiv ()}{\cL \nu\cdot}
\\
\goodType{\Gamma}{\cC (\nu^+ \ \sigma \ \{f\} \ T) \equiv 
    \Sigma (\cC \sigma) T["sf" \ f]}{}
\\\\ \goodTerm{\Gamma}{\cCt \ (\mu^+ \ o \ \{f\}\ t) \equiv ((\cCt \ o), t["sf" \ f][(id, \cCt \ o)]) }{}
}
$$
% $$
% \Rule
% { \goodSeal{\Gamma}{f}{\sigma}{A}
% \\ \goodType{\Gamma, A}{T}{}
% \\ \goodTerm{\Gamma}{t}{T}
% }{}
% $$

@@

\begin{figure}[H]

  \judgebox{[\goodSeal{\Gamma}{s}{\sigma}{A}] 
  = [\goodTerm{\Gamma, \cC\sigma}{s}{A[\pi_1]}] }
  $$
  \Rule[name=Seal Subst]
  {\goodSeal{\Delta}{s}{\sigma}{A}
    \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
  {\goodSeal{\Gamma}{s[\gamma]}{\sigma[\gamma]}{A[\gamma]}}
  =
  \frac
  {\goodTerm{\Delta, \cC\sigma}{s}{A[\pi_1]}
    \quad  \goodSub{\Gamma}{\gamma}{\Delta}  }
  {\goodTerm{\Gamma,\cC\sigma[\gamma]}{s[\gamma^\uparrow]}{A[\pi_1][\gamma^\uparrow]\equiv A[\gamma][\pi_1]}}
  $$
  $$
  \Rule[name=Seal-Id]
  {}
  {\goodSeal{\Gamma}{id_s}{\sigma}{\cC \sigma}}
  = \goodTerm{\Gamma, \cC\sigma}{\pi_2}{\cC\sigma[\pi_1]}
  $$

@@

And these are also the cases we don't need to seal anything (i.e. doing exhaustiveness checking when using recursor), and thus the "Seal-Id" is used. For notation clearance, we will sometime use $"sf'"$ to emphasize that we consider that seal $s$ as a term $("sf'" \ s)$.
